Bioverse: Origins of Life

### 2024-09-06 src/tex/ms.tex
move some old intro stuff into caveats section

The association of chemical pathways to life and planetary environmental conditions offers a new opportunity to test alternate scenarios for life emergence based on planetary-level data collected from the upcoming observations of populations of exoplanets.
Deep-sea hydrothermal scenarios require planetary conditions that may not be met on ocean worlds with large amounts of water, where the water pressure on the ocean floor is high enough to form high-pressure ices~\citep{Noack2016}.
In this case, a testable prediction would be that planets with high-pressure ices do not show biosignatures.
Likewise, if UV light is required to get life started, then there is a minimum planetary UV flux requirement to have an inhabited world.
This requirement is set by competitor thermal processes; if the photoreaction does not move forward at a rate faster than the competitor thermal process(es), then the abiogenesis scenario cannot function.
On the other hand, abundant UV light vastly in excess of this threshold does not increase the probability of abiogenesis, since once the UV photochemistry is no longer limiting, some other thermal process in the reaction network will be the rate-limiting process instead.
Therefore, a putative dependence of life on UV light is best encoded as a step function (see, e.g., Ranjan et al. 2017c; Rimmer et al. 2018; Rimmer, Ranjan \& Rugheimer 2021).

@@

\subsection{Origins of Life Scenarios and their Predictions}\label{sec:predictions}
\todo[inline]{Present widely discussed OOL scenarios and their predictions on exoplanet observables; derive testable hypotheses.}
\label{sec:ool_scenarios}
In this section, we present some of the most prominent origins of life scenarios and their observational predictions.
We focus on the necessary environmental conditions for the processes and reactions inherent to each scenario, and aim to identify distinct observables that are accessible via present and near-future remote sensing techniques.

%\subsection{Hydrothermal vents}
A widely regarded origins-of-life scenario is that abiogenesis happens in hydrothermal vents~\citep[e.g.,][]{Russel2010}.
...
The hydrothermal vents scenario requires a direct contact of an ocean and the planetary mantle/crust.
This requirement is not met on an ocean world with large amounts of water, where the water pressure on the ocean floor is high enough to form high-pressure ices (Noack+2016). % TODO Sleep et al., 2011? Sobotta et al., 2020?
\todo[inline]{see discussion in Kite \& Ford 2018 Sect. 6.4}
\todo[inline]{SR: The sealing away of the planetary interior from the ocean due to high-pressure ice layers is a common assumption for water world exoplanets (in addition to the references above, see e.g. Hu et al. 2021). I'm not convinced it is correct, because of relatively recent evidence showing the possibility of molecular assimitation into such ices and subsequent transport, e.g., https://iopscience.iop.org/article/10.1088/0004-637X/769/1/29/meta, https://iopscience.iop.org/article/10.3847/1538-4357/acb49a/meta, https://iopscience.iop.org/article/10.3847/1538-4357/aa5cfe/meta (I'm sure there are other workers in this area, this is just the group with which I am familiar). Exoplaneteers mostly uniformly accept this proposition, so it's not an unreasonable assumption if you want to run with it so long as its acknowledged and caveated reasonably; I'm just highlighting this for your attention so that you can make an informed decision. }

\textbf{Prediction:} Planets with high-pressure ices do not show biosignatures.

%\subsection{Subaerial ponds}
A different scenario is the emergence of life in hot springs or ponds that are exposed to the planet's atmosphere~\citep[e.g.,][]{Deamer2019}.
...
By its nature, the subaerial ponds scenario relies on rock surfaces exposed to the planetary atmosphere.
Water worlds that have their entire planetary surface covered with water contradict this requirement and do not allow for the wet-dry cycling inherent to this origin of life scenario.
The competition of tectonic stress with gravitational crustal spreading (Melosh 2011) sets the maximum possible height of mountains, which in the solar system does not exceed $\sim \SI{20}{\kilo\meter}$.
Such mountains will be permanently under water on water worlds.
Another impediment to wet-dry cycles is tidal locking of the planet as it stalls stellar tide-induced water movement and diurnal irradiation variability.

@@

>>>>

The association of chemical pathways to life and planetary environmental conditions offers a new opportunity to test alternate scenarios for life emergence based on planetary-level data collected from the upcoming observations of populations of exoplanets.
Deep-sea hydrothermal scenarios require planetary conditions that may not be met on ocean worlds with large amounts of water, where the water pressure on the ocean floor is high enough to form high-pressure ices~\citep{Noack2016}.  % TODO Sleep et al., 2011? Sobotta et al., 2020?
\todo[inline]{see discussion in Kite \& Ford 2018 Sect. 6.4}
In this case, a testable prediction would be that planets with high-pressure ices do not show biosignatures.
Likewise, if UV light is required to get life started, then there is a minimum planetary UV flux requirement to have an inhabited world.
This requirement is set by competitor thermal processes; if the photoreaction does not move forward at a rate faster than the competitor thermal process(es), then the abiogenesis scenario cannot function.
On the other hand, abundant UV light vastly in excess of this threshold does not increase the probability of abiogenesis, since once the UV photochemistry is no longer limiting, some other thermal process in the reaction network will be the rate-limiting process instead.
Therefore, a putative dependence of life on UV light is best encoded as a step function (see, e.g., Ranjan et al. 2017c; Rimmer et al. 2018; Rimmer, Ranjan \& Rugheimer 2021).

@@

@@

\subsubsection{Existence of an atmosphere-crust interface}
By its nature, cyanosulfidic scenario relies on rock surfaces exposed to the planetary atmosphere.
Water worlds that have their entire planeary surface covered with water contradict this requireent and do not allow for the wet-dry cycling inherent to this origin of life scenario.
The competition of tectonic stress with gravitational crustal spreading~\citep{Melosh2011} sets the maximum possible height of mountains, which in the solar system does not exceed $\SI{\sim 20}{\kilo\meter}$.
Such mountains will be permanently underwater on water worlds.
Another impediment to wet-dry cycles may be tidal locking of the planet as it stalls stellar tide-induced water movement and diurnal irradiation variability.
